---
title: "Stay safe! Mapping the crime events in Buffalo, NY (2010-2021)"
author: "Wei Liu"
subtitle: "Explore the spatiotemporal pattern of crime events in the City of Buffalo, New York, USA"
---

# Introduction

Crime, as what it always appears to be, is the negative power in our society that tends to trigger multiple problems. With the increasing demand for higher quality of life in the modern world, the general public attaches more attention to the safety of the living environment as one of the basic factors considered. There has been considerable effort extended to exploring and understanding the pattern of various types of crimes. There exists a sub-field, namely, geographic profiling, which is the marginal subject of criminology and geography, looking into the crime events with both the theoretical basics from criminology and spatial methodologies from geography, with the ultimate goal of promoting suggestions for reducing crime rates and maintaining social safety.  
  
In this project, we want to reveal and understand the spatiotemporal distribution of crime incidents in Buffalo, New York. We are going to achieve our objective mainly by visualization of the crime incidents over time as well as over the study region. Besides, we also want to detect the relationship between crime and corresponding sociodemographic attributes in each defined area. The contribution of this project can be three-folded:  
  
&nbsp; &nbsp; &nbsp; &nbsp; Firstly, we analyze the pattern of crime incidents in Buffalo, NY from multiple dimensions through innovative approaches.  
      
&nbsp; &nbsp; &nbsp; &nbsp; Secondly, we further explore the potential mechanism behind the scene which may play a role in the formation of crime incidents.  
      
&nbsp; &nbsp; &nbsp; &nbsp; Thirdly, based on the aforementioned results, we are able to provide useful suggestions for promoting overall social safety.  


# Data and methods

## Data

1 Crime incidents  

&nbsp; &nbsp; &nbsp; &nbsp; The csv file of crime incidents is in the public domain, available from the website of Open Data Portal of the City of Buffalo.

## Methods


[~ 200 words]

Narrative: Clear narrative description of the data sources and methods. Includes data from at least two sources that were integrated / merged in R.

Code: The code associated with the project is well organized and easy to follow. Demonstrates mastery of R graphics and functions.

Data: The underlying data are publicly accessible via the web and downloaded/accessed within the Rmd script. If you want to use your own data, you must make it available on a website (e.g. Figshare) so that others are able to re-run your code.

You can do bullets like this:

* The first most important thing
* The second most important thing
* The third most important thing

You can do numbers like this:

1. The first most important thing
2. The second most important thing
3. The third most important thing

See [http://rmarkdown.rstudio.com/](http://rmarkdown.rstudio.com/) for all the amazing things you can do.


Here's my first code chunk.
```{r}
1+2
```

Load any required packages in a code chunk (you may need to install some packages):

```{r, message=F, warning=F}
library(tidyverse)
library(leaflet)
library(kableExtra)
knitr::opts_chunk$set(cache=TRUE)  # cache the results for quick compiling
```

## Download and clean all required data
```{r}
n=20
data=data.frame(x=runif(n,-180,180),
                y=runif(n,-60,60),
                size = runif(n, 5, 20),
                category = factor(
                  sample(letters[1:5], n, replace = TRUE)
                  ),
                value = rnorm(n))
```

```{r, results='asis'}
data %>% 
  slice(1:10) %>% #show only 1:n rows
  kable(digits=2,align="c")%>% #make table and round to two digits
  kable_styling(bootstrap_options = 
                  c("striped", "hover", "condensed", "responsive")) #apply other formatting
```

Add any additional processing steps here.

# Results

[~200 words]

Tables and figures (maps and other graphics) are carefully planned to convey the results of your analysis. Intense exploration and evidence of many trials and failures. The author looked at the data in many different ways before coming to the final presentation of the data.

Show tables, plots, etc. and describe them.

```{r, fig.width=6, fig.height=3, fig.cap="Map of completely random data"}
m <- leaflet(data) %>% 
  addTiles() %>% 
  addCircleMarkers(~x, ~y, radius = ~size,color = ~as.factor(category)) %>% 
  addPopups(~x[2], ~y[2], "Random popup")
m  # a map with the default OSM tile layer
```


```{r}
data %>% 
  ggplot(aes(x=x,y=y,col=category))+
  geom_point()
```

# Conclusions

[~200 words]

Clear summary adequately describing the results and putting them in context. Discussion of further questions and ways to continue investigation.

# References

All sources are cited in a consistent manner
